---
title: "EDA"
author: "Partha S Satpathy"
date: "April 24, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r library, echo=TRUE}
library(fma)
library(psych)
library(dplyr)
library(lubridate)
```

```{r readData,echo=TRUE}
path = 'C:/Users/parth/Desktop/Git/Projects/DengAI/Data/'
train_df <- read.csv(paste(path,'dengue_features_train.csv',sep = ""),stringsAsFactors = F)
print(c(nrow(train_df),ncol(train_df)))
head(train_df)

test_df <- read.csv(paste(path,'dengue_features_test.csv',sep = ""),stringsAsFactors = F)
print(c(nrow(test_df),ncol(test_df)))
head(test_df)

combine_df <- rbind(train_df,test_df)
print(c(nrow(combine_df),ncol(combine_df)))
head(combine_df)
```
```{r summaryOfData,echo=TRUE}
describe(combine_df)
```

## The summary shows that the data has a lot of NA values. Let's figure that out.

```{r checkNA,echo=TRUE}
NA_df <- is.na(combine_df)
cols <- colnames(combine_df)
#head(NA_df)
for(j in 1:ncol(NA_df)){
  print(paste(cols[j],'::',sep = ""))
  rowNA <- vector(mode="numeric", length=0)
  for(i in 1:nrow(NA_df)){
    if(isTRUE(NA_df[i,j])){
      rowNA <- c(rowNA,i)
    }
  }
  print(paste('Total null values: ',length(rowNA),sep = ""))
  print("Null values at rows:")
  print(rowNA)
}

```

From the above Analysis it is clear that first 4 columns: city, year, weekofyear, week_start_date do not have NAs. While rest 20 columns have NAs.
Let's first Analyse the Non-NA columns as they will be helpful in imputing the NULL values for other columns.

```{r analyseCity,echo=TRUE}
u = unique(combine_df['city'])
print("No of Unique cities: ")
print(u)
##We have two cities: sj and iq. Let's make this a factor.
combine_df$city = as.factor(combine_df$city)

##Let's check the distribution of the cities
counts <- table(combine_df$city)
print(counts)
barplot(counts, main="City Distribution", xlab="Cities",ylab="No of records")

##The above barchart shows that we have more data on sj city than iq data.
```

As we have two states, let's analyze further for the two cities separately.

```{r analyseYear,echo=TRUE}
analyseByYear <- function(temp_df,city){
  u = unique(temp_df$year)
  print(paste("city ",city,"::No of Unique Years: ",sep = ""))
  print(u)
  
  ##Let's check the distribution of the cities
  counts <- table(temp_df$year)
  print(counts)
  barplot(counts, main=paste("Year Distribution of city ",city,sep = ""), xlab="Years",ylab="No of Weeks")
}
cities = c('sj','iq')
for(i in cities){
  temp_df <- combine_df %>% filter(city==i)
  analyseByYear(temp_df,i)
}

```

So city sj has records from 1990 to 2013. For 1990 it has 35 weeks data and for 2013- 17 weeks. For all other years we have all 52 weeks data.
For city iq- we have data from 2000 to 2013. 2000 has 26 and 2013 has 26 weeks. All others have 52 weeks data.

Let's find out from which week the data starts and ends for the two cities.
```{r analyseByweekofyear,echo=TRUE}
printWeeks <- function(ct,startYear,endYear){
  temp_df <- combine_df %>% filter(city == ct & year == startYear)
  print(paste("STarting Year Weeks for city ",ct,":",sep = ""))
  print(temp_df$weekofyear)
  
  temp_df <- combine_df %>% filter(city == ct & year == endYear)
  print(paste("Ending Year Weeks for city ",ct,":",sep = ""))
  print(temp_df$weekofyear)
}

printWeeks('sj',1990,2013)
printWeeks('iq',2000,2013)
```

Let's check that the week start day is correct for all the rows.

```{r checkWeekSTartDay,echo=TRUE}
calcWeek <- as.list(lubridate::week(ymd(combine_df$week_start_date)))
actualweek <- as.list(combine_df$weekofyear)
length(setdiff(calcWeek,actualweek))
```
So, the week start day is correct for all rows.

Now, let's start imputing NA values.
We will start with ndvi_ne

```{r imputendvi_ne,echo=TRUE}
imputendvi_ne <- function(temp_df,city,startYear,startWeek,endWeek){
  ndvi_ne_ts <- ts(temp_df$ndvi_ne,start=c(startYear,startWeek),end=c(2013,endWeek),frequency = 52)  
  print(paste(city," : ",sep = ""))
  #print(ndvi_ne_ts)
  ts.plot(ndvi_ne_ts,col=2)
}

imputendvi_ne((combine_df %>% filter(city == 'sj')),'sj',1990,18,17)
imputendvi_ne((combine_df %>% filter(city == 'sj')),'iq',2001,1,26)
```

